---
title: "IMDB(2006-2016)"
author: "Ambuj Bhardwaj"
date: "19 April 2019"
output: 
  html_document:
    toc: true
    code_folding: hide      
---

````{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
````

#Introduction

Dataset available at <https://www.kaggle.com/PromptCloudHQ/imdb-data>

We will be using *IMDB-Movie-Data.csv* for this project.

```{r, include=TRUE}
imdb <- read.csv(file = "../input/IMDB-Movie-Data.csv", header = TRUE, stringsAsFactors = FALSE)
```
#Data Exploration

Before we start manipulating the data, first we will check the structure of our dataset.

```{r}
str(imdb)
```

From above output we can see that there are 1000 observations or row and 12 variables or columns in imdb dataframe that we created in our first code block.

Let us check the column names in our dataset
```{r}
colnames(imdb)
```

Here two column names at index 8 and 11 looks too long. Let's rename these two.

```{r}
colnames(imdb)[c(8,11)] <- c("Runtime", "Revenue")
colnames(imdb)
```


```{r, warning= FALSE, message= FALSE}
#load libraries that we will use in our EDA process
library(ggplot2)
library(d3Tree)
library(dplyr)
library(wordcloud)
```

##Exploration by Year

Using d3Tree package to group the movies by their year of release to form an interactive branching relationship plot.

```{r}
yt <- imdb %>% group_by(Year) %>% arrange(Year) %>% select(Year, Title) %>% count(Year, Title)
d3tree(list(root = df2tree(rootname = 'IMDB',struct = as.data.frame(yt)), layout = 'collapse'))
```

Using ggplot2 package to generate barplots to get more depth.

```{r}
ggplot(data = imdb, aes(x = factor(Year))) + geom_bar(stat = 'count') + geom_text(stat = 'count', aes(label = ..count..), vjust = -1) + theme(panel.background = element_rect('transparent')) + geom_hline(yintercept = c(0,100,200,300)) + xlab("Year") + ylab("Number of Movies")
```

Lets have a glance at the Revenue of Movies each year

```{r}
newimdb<-imdb
newimdb$Revenue[which(is.na(newimdb$Revenue))]<-0 #cleaning NA values in Revenue
ggplot(data = newimdb, aes(x = factor(Year), y = Revenue, fill = factor(Year)))+ geom_violin() + xlab("Year") + theme_bw()
```

```{r}
ggplot(data = newimdb, aes(x = factor(Year), y = Revenue)) + geom_boxplot() + xlab("Year")
```


```{r}
year_revenue<-aggregate(Revenue~Year, data = newimdb , sum)
ggplot(data = year_revenue, aes(factor(Year), Revenue, fill = factor(Year))) + 
  geom_bar(stat = 'identity') +
  geom_text(aes(label = Revenue), vjust = -0.75, cex = 2.5) + 
  xlab("Year") + ylab("Revenue in Millions")
```

##Exploration by Rating

Let us also see how the ratings affect the Revenue of Movies. But first we must do some processing to club the ratings into intervals.

```{r}
rr<- aggregate(Revenue~Rating, data = newimdb, sum)
rr$Rating_Interval<- cut(rr$Rating, breaks = seq(from = 1, to = 10, by = 1))
head(rr)
```
Now we can create a boxplot to see how the ratings have affected the Revenue.

```{r}
ggplot(data = rr, aes(x = factor(Rating_Interval), y = Revenue, fill = Rating_Interval)) + geom_boxplot() + xlab("Ratings")
```

To further drilldown into this we must also see how many movies were categorised in all interval ranges
```{r}
newimdb$ri<- cut(newimdb$Rating, breaks = seq(1,10, by =1))
rc <- newimdb %>% group_by(ri) %>% count(ri)
ggplot(data =  rc, aes(x = ri, y = n)) + geom_bar(stat = 'identity') + 
  xlab("Rating") + ylab("Number of MOvies")
```

Now it is clear that maximum number of movie ratings ranges between 6-8. Hence the Revenue is more for the movies in this range of ratings. This clearly indicates that Revenues is directly proportional to the Ratings received.

##Exploration by Genre

Let us now check the data with respect to Genre of movies

```{r}
#gets unique words in all genre column
w<-unique(unlist(strsplit(unique(newimdb$Genre), ",")))

#UDF for counting the appearance of unique words
wc<- function(x){
  length(grep(x, newimdb$Genre))
}

genrecount<- as.data.frame(sapply(w,wc))
genrecount$genre <- rownames(genrecount)
colnames(genrecount)<- c("count","genre")

#Arranging in ascending order by count
genrecount<- arrange(genrecount, count)

#plotting bar grapgh
ggplot(data = genrecount, aes(x = genre, y = count)) + 
  geom_bar(stat = 'identity') + 
  coord_flip()
```

From above bar graph it is clear that Drama and Action movies are fairly popular, followed by Comedy and Adventure.


```{r}
#UDF for counting Revenue distribution by Genre
RevSum<- function(x){
  sum(newimdb[(grep(x,newimdb$Genre)), "Revenue"])  
}

rev<-as.data.frame(sapply(w,RevSum))
rev$genre <- rownames(rev)
colnames(rev)[1] <- "Revenue"


ggplot(data = rev, aes(x = genre, y = Revenue)) + 
  geom_bar(stat = 'identity') +
  xlab("Genre") + ylab("Revenue in Millions") +
  coord_flip()
```

Above grapgh clearly visualises the distribution of Revenue by Genre of the movie.

#Conclusion
```{r}
wordcloud(genrecount$genre, genrecount$count, scale = c(3,0.5))
```

1. Positive trend in number of movie release every year.
2. Revenue of movies is increasing YoY.
3. Maximum number of movies were rated between 6-8.
4. Drama movies tends to dominate the genre.
5. Action and Adventure movies bring most Revenue.


